---
title: "Analysis"
output: html_notebook
editor_options:
  chunk_output_type: inline
---

# Analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "#>",
                      collapse = TRUE)
knitr::opts_knit$set(root.dir = '../')
```

## Setup

Load reqired packages

```{r load packages for data processing, hide=TRUE}
#data wrangling and visualization
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(ggforce)
library(ggridges)
library(ggsci)
library(wesanderson)
library(gghighlight)
library(GGally)
library(tidygraph)
library(ggraph)
library(graphlayouts)

#network analysis
library(network)
library(igraph)

#GLMM
library(lme4)
library(car)


```
```{r load packages for data analysis, hide=TRUE}
library(lme4)
library(MuMIn)
library(car)
```

## Data overview

load relatedness results

```{r import relatedness results}
load("data/relate_results.RData")
load("data/sample_Info.RData")
load("data/data_overview.RData")

```

```{r}
str(relate_results)
```




load association data
```{r import association data}
load("data/co-roosting-network.RData")
str(network)

```



## network of relatedness
let's try to visualize the relatedness relationships with network, in which each individual is a node and edges show genetic relationship

```{r}
#select relatedness data (ngsrelate rab) from mpala
#relatedness_rab <- relate_results%>% subset(a_Group %in% c("mpala", "mpala_chicks") & b_Group %in% c("mpala", "mpala_chicks")) %>% select(ID_a, ID_b, rab)

relatedness_rab <- relate_results %>% select(ID_a, ID_b, rab)

#create node list
ID_a <- relatedness_rab %>% distinct(ID_a) %>% rename(label = ID_a)

ID_b <- relatedness_rab %>% distinct(ID_b) %>% rename(label = ID_b)

nodes <- full_join(ID_a, ID_b, by = "label")

nodes <- nodes %>% rowid_to_column("id")

nodes <- left_join(nodes, sample_info, by = c("label"= "ID"))

nodes

```


```{r}
#create edge list
edges <- relatedness_rab %>% left_join(nodes, by = c("ID_a" = "label")) %>% rename(id_a = id)

edges <- edges %>% left_join(nodes, by = c("ID_b" = "label")) %>% rename(id_b = id) %>% rename (weight = rab)

edges <- select(edges, id_a, id_b, weight)

edges
```


using igraph

tidygraph and ggraph


```{r}
#rab_tidy <- tbl_graph(nodes = nodes, edges = edges, directed = FALSE)

#theme_set(dark_theme_classic())

rab_igraph <- graph_from_data_frame(d = edges, vertices = nodes, directed = FALSE)

#create a cutoff under which edges are not displayed or manually set cutoff to 0.1
cut.off <- mean(edges$weight) 
net.sp <- delete_edges(rab_igraph, E(rab_igraph)[weight<0.1])

#generate colours based on sex
V(net.sp)$shape <- case_when(V(net.sp)$sex == "f" ~ "circle",
                                V(net.sp)$sex == "m" ~ "triangle")

Ten_colors <-  scales::viridis_pal()(10)

#generate shape based on group
V(net.sp)$color <- case_when(V(net.sp)$Group_ID == "mpala" ~ Ten_colors[1],
                                V(net.sp)$Group_ID == "mpala_chicks" ~ Ten_colors[2],
                                V(net.sp)$Group_ID == "wing_tags" ~ Ten_colors[3],
                                V(net.sp)$Group_ID == "5509" ~ Ten_colors[4],
                                V(net.sp)$Group_ID == "5931-5938" ~ Ten_colors[5],
                                V(net.sp)$Group_ID == "dump" ~ Ten_colors[6],
                                V(net.sp)$Group_ID == "wt025" ~ Ten_colors[7],
                                 V(net.sp)$Group_ID == "5512" ~ Ten_colors[8],
                                 V(net.sp)$Group_ID == "RRWB" ~ Ten_colors[9],
                                 V(net.sp)$Group_ID == "ROOP" ~ Ten_colors[10]
                             )

ggraph(net.sp, layout = "stress",bbox = 50) + 
 # geom_node_point() +
  geom_edge_link2(aes(width = weight), alpha = 0.2) + 
  scale_edge_width(range = c(0, 3)) +
  geom_node_point(color=V(net.sp)$color, shape = V(net.sp)$shape, size= 2)  +
  labs(edge_width = "rab", color = "Group_ID", shape = "sex") +
  theme_graph()+
  theme(legend.position = "bottom") 




```

```{r}
group <- c("mpala" , "mpala_chicks","wing_tags","5509" , "5931-5938", "dump" ,"wt025","5512" ,"RRWB","ROOP")
ID <- (1:10)
sex <- c("male" , "female","male", "female","male", "female","male", "female","male", "female" )

test <- as.data.frame(cbind(group, ID, sex))

ten_colors_sorted <- c(Ten_colors[4],Ten_colors[8],Ten_colors[5],Ten_colors[6],Ten_colors[1],Ten_colors[2],Ten_colors[10],Ten_colors[9], Ten_colors[3], Ten_colors[7])

theme_set(theme_classic())

ggplot(test, aes(x=group, y=sex)) +
  
  geom_point(aes(color=group, shape = sex)) +
  
  scale_color_manual(values = ten_colors_sorted)
  
  

```





```{r}
# Community detection (by optimizing modularity over partitions):
clp <- cluster_optimal(net.sp)
class(clp)


V(net.sp)$community <- clp$membership

V(net.sp)$community2 <- case_when(V(net.sp)$community == "3" ~ Ten_colors[1],
                                V(net.sp)$community == "4" ~ Ten_colors[3],
                                V(net.sp)$community == "2" ~ Ten_colors[7],
                                V(net.sp)$community == "1" ~ Ten_colors[2],
                                V(net.sp)$community == "5" ~ Ten_colors[10],
                                V(net.sp)$community == "6" ~ Ten_colors[9],
                                V(net.sp)$community == "7" ~ "brown")


ggraph(net.sp, layout = "stress",bbox = 50) + 
  
  geom_edge_link(aes(width = weight), alpha = 0.2) + 
  scale_edge_width(range = c(0, 3)) +
  geom_node_point(color=V(net.sp)$community2, shape = V(net.sp)$shape, size= 5) +
  geom_node_point(color=V(net.sp)$color, shape = V(net.sp)$shape, size= 3) +
  geom_node_text(aes(filter = (label == "  "), label = label, size = degree(net.sp))) +
  labs(edge_width = "rab", color = "Group_ID", shape = "sex") +
  theme_graph()+
  theme(legend.position = "bottom") 

```


try better visualization with "backbone" following the tutorial http://mr.schochastics.net/netVizR.html

```{r}
bb <- layout_as_backbone(rab_igraph,keep = 0.1)
E(rab_igraph)$col <- FALSE
E(rab_igraph)$col[bb$backbone] <- TRUE

ggraph(rab_igraph,layout = "manual",x = bb$xy[,1],y = bb$xy[,2])+
  

  geom_edge_link(aes(width = weight), alpha = 0.2) + 
  scale_edge_width(range = c(0, 3)) +
  
  geom_node_point(aes(fill = Group_ID),shape = 21)+
  scale_fill_brewer(palette = "Set1")+
  theme_graph()+
  theme(legend.position = "none")



```


#association network

extract the networks for the six different seasons

```{r}
#get long format of associations from network data

for(i in 1:6) {
  
 assoc_network <- network[[i]] %>% as.data.frame %>% tibble::rownames_to_column() %>% tidyr::pivot_longer(-rowname) %>% filter(rowname != name)

names(assoc_network) <- c("id_a","id_b",paste("assoc", i, sep = "_")) 

assign(paste("assoc_nw", i, sep = "_"), assoc_network)
}

#assoc_nw_1 <- network[[1]] %>% as.data.frame %>% tibble::rownames_to_column() %>% tidyr::pivot_longer(-rowname) %>% filter(rowname != name)

#names(assoc_nw_1) <- c("id_a","id_b","assoc_1") 


#combine association data in one table

assoc_nw <- full_join(assoc_nw_1, assoc_nw_2, by = c("id_a", "id_b"))

assoc_nw <- full_join(assoc_nw, assoc_nw_3, by = c("id_a", "id_b"))

assoc_nw <- full_join(assoc_nw, assoc_nw_4, by = c("id_a", "id_b"))

assoc_nw <- full_join(assoc_nw, assoc_nw_5, by = c("id_a", "id_b"))

assoc_nw <- full_join(assoc_nw, assoc_nw_6, by = c("id_a", "id_b"))

```

#statistical analysis: GLMM
Run general and generalized linear mixed models with the package `lme4`.

Model formular to test influence on relatedness: is relatedness influenced by group membership and is this different for males and females?

response: dyadic relatedness
main predictors: association and sex
fixed control. group membership
random intercept: identity
coverage?
need to run for every relatedness estimator
for association: test 


## Check the data

```{r}
str(relate_results)
```

inspect the data:
inspect the distribution of the predictor 
```{r}
hist()
```

and the responses
```{r}
hist()
```

```{r}
hist()
```

is there a need to transform any of the variables?

Z-transform all covariates to a mean of 0 and a sd of 1 prior to fitting the model to allow easier interpretation of the estimates. But still need to inform about original means and sd.

```{r z-transformation}

relate_results$rab_z <- scale(relate_results$rab)
relate_results$pi_HAT_z <- scale(relate_results$pi_HAT) 
relate_results$Normalized2AlleleDifference_z <- scale(relate_results$Normalized2AlleleDifference)
relate_results$coverage_z <- scale(relate_results$coverage)
relate_results$nbSNP_z <- scale(relate_results$nbSNP)
relate_results$nSites_z <- scale(relate_results$nSites)

```

check correlations among the predictors

```{r}
pred.vars <- c()

vars.cor <- cor(data[,pred.vars])

max(abs(vars.cor[upper.tri(vars.cor)]))
```


## fit the model

full model
```{r full glmm}
res <- lm(response ~ predictor1 + predictor2, data)
```

for hypothesis testing, we also need a null model
```{r null glmm}

```

## model diagnostics

### check of assumptions

normality of residuals

```{r}
residuals(res)

hist(residuals(res), probability = T)

x <- seq(from = min(residuals(res)), to = max(residuals(res)), length.out = 100)

lines(x = x, y = dnorm(x, mean = 0, sd = sd(residuals(res))))
```

qqplot
```{r}
#in an indeal case, all points would fall on a straight line
qqnorm(residuals(res))
qqline(residuals(res))
```

homogeneity of residuals

```{r}
#no obvious pattern should be visible
fitted(res)

plot <- (x = fitted(res), y = residuals(res), pch = 19)
```

### model stability


fitted values (DFFit)

```{r}
dffits(res) #standardized DFFit-values

max(abs(dffits(res))) #absolute value >~2 is a reason to worry

hist(dffits(res))
```

model estimates (DFBeta)

```{r}
dfbeta(res) #unstandardized DFBeta

xx <- cbind(coefficients(res), coefficients(res) +
              t(apply(X = dfbeta(res), MARGIN = 2, FUN = range)))

colnames(xx) <- c("orig", "min", "max")
round(xx, 5)
```

```{r}
#standardized DFBeta
dfbetas(res) #absolute values larger than 1 or 2 are a reason to worry
```

Cook's distance
```{r}
max(cooks.distance(res)) #check recommended thresholds in Mundry handout 02b
```

leverage
```{r}
max(as.vector(influence(res)$hat))
```

Variance Inflation Factors (VIF) requires `car` package.

```{r}
vif(res) #check recommended thresholds in Mundry handout 03b
```


plot various diagnostics
SEE FUNCTION IN MUNDRY SCRIPT 02b



## Inference

Protection agains multiple testing with full-null model comparison


Tests of the individual predictors


Measuring effect isze of individual predictors






vgl. Dal Pesco et al. 2021

Check for need of random slopes




Model formular to test influence on relatedness

response: dyadic relatedness
main predictor: association and sex
fixed control. group membership
random intercept: identity
coverage?
need to run for every relatedness estimator
for association: test 

lmer(relatedness ~ association + group + sex + (1|ID_a) + (1|ID_b), data = relate_results, REML = F)

testing this
```{r}
model_simpletest <- lmer(rab_z ~ dyad_group + dyad_sex + (1|ID_a) + (1|ID_b), data = relate_results, REML = F)
model_simpletest
```




















